---
title: "ARG-NLI Data Analysis Hard Threshold"
output: html_notebook
---
```{r}
library(lme4)
library(broom.mixed)
library(dplyr)
```

# Data Loading and Preparation

In this section, we load in the data, transform all relevant character vectors to factors, and clean up NAs in the data.

```{r read Data}
hard_threshold <- read.csv("D:\\OneDrive - University of Dundee\\PhD\\ARG-NLI_project\\feature_analysis\\automatic\\new_features\\corpus_threshold_hard_0.999_standardised_and_log_scaled.csv")

```

```{r transform character vectors to factors}
factor_vars <- c(
  "seg1_has_deduction", "seg2_has_deduction",
  "seg1_has_refutation", "seg2_has_refutation", 
  "seg1_has_condition", "seg2_has_condition",
  "seg1_has_explanation", "seg2_has_explanation", 
  "seg1_has_adverb_of_emphasis", "seg2_has_adverb_of_emphasis",
  "seg1_has_justification", "seg2_has_justification", 
  "seg1_has_modal_verbs", "seg2_has_modal_verbs",
  "seg1_senti", "seg2_senti", "CorpusID"
)

hard_threshold[factor_vars] <- lapply(hard_threshold[factor_vars], function(x) as.factor(as.character(x)))
```


```{r define outcomes}
is_entailment <- "is_entailment"
is_contradiction <- "is_contradiction"
is_neutral <- "is_neutral"
is_support <- "is_support"
is_attack <- "is_attack"
support_is_entailment <- "arg_nli_support"
attack_is_contradiction <- "arg_nli_attack"

outcome <- is_entailment
```


```{r clean up NAs}
# Drop rows with any missing values
hard_threshold_complete <- hard_threshold[complete.cases(hard_threshold), ]
```

# Step 1: Block model fitting
In the first step, we split up the entire set of features into 8 linguistically motivated blocks and fit one model for each block that includes these features plus the number of tokens as a baseline feature. The gain in AIC is then compared for each block with the baseline model with the number of tokens as the only feature. Since we cannot guarantee that there is no interdependence introduced by the different corpora, the corpus is always used as a random effect.

```{r define blocks}
blocks <- list(
  lexical_markers = c("seg1_has_deduction", "seg2_has_deduction",
                      "seg1_has_refutation", "seg2_has_refutation", 
                      "seg1_has_condition", "seg2_has_condition",
                      "seg1_has_explanation", "seg2_has_explanation", 
                      "seg1_has_adverb_of_emphasis", "seg2_has_adverb_of_emphasis",
                      "seg1_has_justification", "seg2_has_justification", 
                      "seg1_has_modal_verbs", "seg2_has_modal_verbs"),
  lexical_sim = c("tfidf_cosine_similarity"),
  semantic = c("cosine_similarity"),
  structure = c("seg1_sentence_length_mean", "seg2_sentence_length_mean"),
  #                     "seg1_n_tokens", "seg2_n_tokens"),
  syntactic = c("seg1_dependency_distance_mean", "seg2_dependency_distance_mean",
                "seg1_prop_adjacent_dependency_relation_mean", "seg2_prop_adjacent_dependency_relation_mean",
                "seg1_avg_subclauses", "seg2_avg_subclauses",
                "seg1_avg_tree_depth", "seg2_avg_tree_depth"),
  perplexity = c("seg1_perplexity", "seg2_perplexity", #"seg1_given_seg2_perplexity", 
                 "seg2_given_seg1_perplexity"),
  indirectness = c("seg1_impScore", "seg2_impScore", "prag_distance"),
  sentiment = c("seg1_senti", "seg2_senti")
)

interactions <- list(
  c("seg1_avg_subclauses", "seg1_sentence_length_mean"),
  c("seg2_avg_subclauses", "seg2_sentence_length_mean"),
  c("seg1_avg_tree_depth", "seg1_avg_subclauses"),
  c("seg2_avg_tree_depth", "seg2_avg_subclauses"),
  c("seg1_avg_tree_depth", "seg1_sentence_length_mean"),
  c("seg2_avg_tree_depth", "seg2_sentence_length_mean"),
  c("seg1_dependency_distance_mean", "seg1_sentence_length_mean"),
  c("seg2_sentence_length_mean", "seg2_dependency_distance_mean"),
  c("seg1_dependency_distance_mean", "seg1_avg_subclauses"),
  c("prag_distance", "cosine_similarity"),
  c("tfidf_cosine_similarity", "cosine_similarity"),
  c("seg2_dependency_distance_mean", "seg2_avg_subclauses"),
  c("seg2_sentence_length_mean", "seg2_n_tokens"),
  c("seg2_avg_subclauses", "seg2_n_tokens"),
  c("seg1_avg_subclauses", "seg1_n_tokens"),
  c("seg1_sentence_length_mean", "seg1_n_tokens"),
  c("seg2_dependency_distance_mean", "seg2_n_tokens"),
  c("tfidf_cosine_similarity", "prag_distance"),
  c("seg1_avg_tree_depth", "seg1_dependency_distance_mean"),
  c("seg2_avg_tree_depth", "seg2_n_tokens"),
  c("seg1_dependency_distance_mean", "seg1_n_tokens"),
  c("seg1_avg_tree_depth", "seg1_n_tokens"),
  c("seg2_dependency_distance_mean", "seg2_prop_adjacent_dependency_relation_mean"),
  c("seg1_prop_adjacent_dependency_relation_mean", "seg1_dependency_distance_mean")
)

```

```{r helper variables to formula}
vars_to_formula <- function(vars, interactions = NULL) {
  f <- paste(vars, collapse = " + ")
  if (!is.null(interactions) && length (interactions) > 0) {
    int_terms <- sapply(interactions, function(x) paste(x, collapse = ":"))
  }
  f
}
```

```{r block fitting function}
fit_block_models_safe <- function(df, outcome, blocks, random = "(1 | CorpusID)",
                                  base_controls = c("seg1_n_tokens", "seg2_n_tokens"),
                                  family = binomial) {
  library(lme4)
  library(performance)
  library(dplyr)
  
  models <- list()
  results <- tibble()
  
  # baseline model
  base_formula <- paste0(outcome, " ~ ", paste(base_controls, collapse = " + "), " + ", random)
  message("Fitting baseline model: ", base_formula)
  m_base <- glmer(as.formula(base_formula), data = df,
                  family = family, control = glmerControl(optimizer = "bobyqa"))
  models[["baseline"]] <- m_base
  results <- bind_rows(results, tibble(
    block = "baseline", AIC = AIC(m_base),
    delta_AIC = NA, LRT = NA, p = NA, r2_marg = NA
  ))
  
  # loop over blocks
  for (b in names(blocks)) {
    vars <- c(base_controls, blocks[[b]])
    fstring <- paste0(outcome, " ~ ", paste(vars, collapse = " + "), " + ", random)
    message("Fitting block-only model: ", b)
    
    m <- tryCatch(
      glmer(as.formula(fstring), data = df, family = family,
            control = glmerControl(optimizer = "bobyqa")),
      error = function(e) { warning("Model failed for block ", b, ": ", e$message); return(NULL) }
    )
    
    if (!is.null(m)) {
      models[[b]] <- m
      comp <- tryCatch(anova(m_base, m, test = "Chisq"), error = function(e) NULL)
      r2 <- tryCatch(performance::r2_nakagawa(m)$R2_marginal, error = function(e) NA)
      
      row <- tibble(
        block = b,
        AIC = AIC(m),
        delta_AIC = if (!is.null(comp)) AIC(m_base) - AIC(m) else NA,
        LRT = if (!is.null(comp)) comp$Chisq[2] else NA,
        p = if (!is.null(comp)) comp$`Pr(>Chisq)`[2] else NA,
        r2_marg = r2
      )
      results <- bind_rows(results, row)
    } else {
      results <- bind_rows(results, tibble(
        block = b, AIC = NA, delta_AIC = NA, LRT = NA, p = NA, r2_marg = NA
      ))
    }
  }
  
  list(models = models, summary = results)
}
```


```{r fit block models}
block_results_entailment <- fit_block_models_safe(hard_threshold_complete, outcome = outcome, blocks = blocks)

block_results$summary %>%
  arrange(AIC)  # smaller AIC = better fit
```

```{r plot block models}
library(ggplot2)
ggplot(block_results$summary[-1,], aes(x = reorder(block, -delta_AIC), y = delta_AIC)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(x = "Feature Block", y = "ΔAIC improvement over baseline",
       title = "Relative contribution of feature blocks")
```

Based on the results from the blockwise fitting, the significant features for each block are identified.

```{r}
extract_block_details <- function(block_results) {
  library(broom.mixed)
  lapply(names(block_results$models), function(b) {
    m <- block_results$models[[b]]
    if (is.null(m)) return(NULL)
    broom.mixed::tidy(m, effects = "fixed", conf.int = TRUE, exponentiate = TRUE) %>%
      mutate(block = b)
  }) %>% bind_rows() %>% filter(term != "(Intercept)")
}
```

```{r}
block_details <- extract_block_details(block_results)
```


```{r}
strong_blocks <- c("semantic", "indirectness", "lexical_sim", "perplexity")

sig_features <- block_details %>%
  filter(block %in% strong_blocks,
         p.value < 0.05,
         abs(estimate - 1) > 0.05) %>%
  arrange(block, p.value) %>%
  select(block, term, estimate, p.value, conf.low, conf.high)

sig_features
```

```{r}
cor(hard_threshold_complete[, unique(sig_features$term)], use = "pairwise.complete.obs")
```
# Step 2: Refine features from strong blocks

In the second step, we fit a model with all singificant features from the strong blocks identified in Step 1 and compare it to a baseline model with only the number of tokens. Then, we sequentially refine the model by adding or removing features, as well as introducing interactions between features that can interact from a theoretical perspective. The best fit is identified by improvements in AIC and R2 in comparison to the former best fit model. In addition, VIF values are checked at every step. Inside the models, ANOVA analysis is used to identify significant features and to remove features with minor or no influence on the model.

```{r}
final_features <- c(
  "cosine_similarity",
  "prag_distance",
  "seg2_given_seg1_perplexity",
  "seg2_n_tokens",
  "seg1_n_tokens",
  "seg1_perplexity",
  "seg2_perplexity",
  "seg1_impScore",
  "seg2_impScore",
  "tfidf_cosine_similarity"
)

f_final <- as.formula(
  paste(outcome, "~", paste(final_features, collapse = " + "), "+ (1 | CorpusID)")
)

model_baseline_entailment <-glmer(is_entailment ~ (1|CorpusID) + seg1_n_tokens + seg2_n_tokens, data = hard_threshold_complete,
                       family = binomial,
                       control = glmerControl(optimizer = "bobyqa"))

model_blocks_entailment <- glmer(f_final, data = hard_threshold_complete,
                     family = binomial,
                     control = glmerControl(optimizer = "bobyqa"))
```


```{r}
model_without_length_entailment <- glmer(is_entailment ~ (1|CorpusID) 
                        + cosine_similarity 
                        + prag_distance
                        + seg2_given_seg1_perplexity 
                        + seg1_perplexity 
                        + seg2_perplexity
                        #+ seg2_n_tokens
                        #+ seg1_n_tokens
                        + seg1_impScore
                        + seg2_impScore
                        + tfidf_cosine_similarity, 
                        data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))
```

```{r}
library(car)

anova(model_baseline_entailment, model_blocks_entailment, model_without_length_entailment, test = "Chisq")
print("R2 baseline")
performance::r2_nakagawa(model_baseline_entailment)
print("R2 all block features")
performance::r2_nakagawa(model_blocks_entailment)
print("R2 without length")
performance::r2_nakagawa(model_without_length_entailment)

vif(model_baseline_entailment)
vif(model_blocks_entailment)
vif(model_without_length_entailment)
#vif(model_interact_with_length)
```
In this step, the Model with all significant features from all blocks is the best model with and R2 conditional = 0.55 and marginal = 0.4.

## Interaction perplexity features

```{r}
model_interact_perplexity_entailment <-  glmer(is_entailment ~ (1|CorpusID) 
                        + cosine_similarity 
                        + prag_distance
                        + seg2_given_seg1_perplexity * seg1_perplexity 
                        + seg2_given_seg1_perplexity * seg2_perplexity
                        + seg2_n_tokens
                        + seg1_n_tokens
                        + seg1_impScore
                        + seg2_impScore
                        + tfidf_cosine_similarity, 
                        data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))
```


```{r}
library(car)

anova(model_blocks_entailment, model_interact_perplexity_entailment, test = "Chisq")

print("R2 all block features")
performance::r2_nakagawa(model_blocks_entailment)
print("R2 with perplexity interactions")
performance::r2_nakagawa(model_interact_perplexity_entailment)

vif(model_blocks_entailment)
vif(model_interact_perplexity_entailment)
#vif(model_interact_with_length)
```
The interaction between perplexity features gives a significant gain in performance Conditional R2: 0.548; Marginal R2: 0.398. 

## Interaction between cosine similarity and tfidf cosine similarity
```{r}
model_interact_tfidf_entailment <- glmer(is_entailment ~ (1|CorpusID) 
                        + cosine_similarity * tfidf_cosine_similarity
                        + prag_distance
                        + seg2_given_seg1_perplexity * seg1_perplexity 
                        + seg2_given_seg1_perplexity * seg2_perplexity
                        + seg2_n_tokens
                        + seg1_n_tokens
                        + seg1_impScore
                        + seg2_impScore
                        #+ tfidf_cosine_similarity 
                        , data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))
```

```{r}
library(car)

anova(model_interact_perplexity_entailment, model_interact_tfidf_entailment, test = "Chisq")

print("R2 with perplexity interactions")
performance::r2_nakagawa(model_interact_perplexity_entailment)

print("R2 add tfidf interaction")
performance::r2_nakagawa(model_interact_tfidf_entailment)

vif(model_interact_perplexity_entailment)
vif(model_interact_tfidf_entailment)
#vif(model_interact_with_length)
```
## Interaction between cosine similarity and pragmatic distance
```{r}
model_interact_prag_entailment <- glmer(is_entailment ~ (1|CorpusID) 
                        + cosine_similarity * tfidf_cosine_similarity
                        + cosine_similarity * prag_distance
                        #+ prag_distance
                        + seg2_given_seg1_perplexity * seg1_perplexity 
                        + seg2_given_seg1_perplexity * seg2_perplexity
                        + seg2_n_tokens
                        + seg1_n_tokens
                        + seg1_impScore
                        + seg2_impScore
                        #+ tfidf_cosine_similarity 
                        , data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))
```

```{r}
library(car)

anova(model_interact_tfidf_entailment, model_interact_prag_entailment, test = "Chisq")

print("R2 add tfidf interaction")
performance::r2_nakagawa(model_interact_tfidf_entailment)

print("R2 with pragmatic distance interaction")
performance::r2_nakagawa(model_interact_prag_entailment)

vif(model_interact_tfidf_entailment)
vif(model_interact_prag_entailment)
#vif(model_interact_with_length)
```

## Intearction conditional perplexity and pragmatic distance
We don't keep the cosine x prag_distance interaction
```{r}
model_interact_prag_per_entailment <- glmer(is_entailment ~ (1|CorpusID) 
                        + cosine_similarity * tfidf_cosine_similarity
                        + cosine_similarity
                        + prag_distance * seg2_given_seg1_perplexity
                        + seg2_given_seg1_perplexity 
                        * seg1_perplexity 
                        + seg2_given_seg1_perplexity 
                        * seg2_perplexity
                        + seg2_n_tokens
                        + seg1_n_tokens
                        + seg1_impScore
                        + seg2_impScore
                        #+ tfidf_cosine_similarity 
                        , data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))
```

```{r}
library(car)

anova(model_interact_tfidf_entailment, model_interact_prag_per_entailment, test = "Chisq")

print("R2 add tfidf interaction")
performance::r2_nakagawa(model_interact_tfidf_entailment)

print("R2 with pragmatic x perplexity distance interaction")
performance::r2_nakagawa(model_interact_prag_per_entailment)

vif(model_interact_tfidf_entailment)
vif(model_interact_prag_per_entailment)
#vif(model_interact_with_length)
```

## Interaction cosine similarity and perplexity
```{r}
model_interact_sem_per_entailment <- glmer(is_entailment ~ (1|CorpusID) 
                        + cosine_similarity * tfidf_cosine_similarity
                        + cosine_similarity * seg2_given_seg1_perplexity
                        + prag_distance 
                        + seg2_given_seg1_perplexity 
                        * seg1_perplexity 
                        + seg2_given_seg1_perplexity 
                        * seg2_perplexity
                        + seg2_n_tokens
                        + seg1_n_tokens
                        + seg1_impScore
                        + seg2_impScore
                        #+ tfidf_cosine_similarity 
                        , data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))
```

```{r}
library(car)

anova(model_interact_tfidf_entailment, model_interact_sem_per_entailment, test = "Chisq")

print("R2 add tfidf interaction")
performance::r2_nakagawa(model_interact_tfidf_entailment)

print("R2 with cosine x perplexity distance interaction")
performance::r2_nakagawa(model_interact_sem_per_entailment)

vif(model_interact_tfidf_entailment)
vif(model_interact_sem_per_entailment)
#vif(model_interact_with_length)
```

## Interaction tfidf x conditional perplexity

```{r}
model_interact_tfidf_per_entailment <- glmer(is_entailment ~ (1|CorpusID) 
                        + cosine_similarity * tfidf_cosine_similarity
                        + tfidf_cosine_similarity * seg2_given_seg1_perplexity
                        + prag_distance 
                        + seg2_given_seg1_perplexity 
                        * seg1_perplexity 
                        + seg2_given_seg1_perplexity 
                        * seg2_perplexity
                        + seg2_n_tokens
                        + seg1_n_tokens
                        + seg1_impScore
                        + seg2_impScore
                        #+ tfidf_cosine_similarity 
                        , data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))
```

```{r}
library(car)

anova(model_interact_tfidf_entailment, model_interact_tfidf_per_entailment, test = "Chisq")

print("R2 add tfidf interaction")
performance::r2_nakagawa(model_interact_tfidf_entailment)

print("R2 with tfidf x perplexity distance interaction")
performance::r2_nakagawa(model_interact_tfidf_per_entailment)

vif(model_interact_tfidf_entailment)
vif(model_interact_tfidf_per_entailment)
#vif(model_interact_with_length)
```



```{r}
model_interact_seg1_len_per_entailment <- glmer(is_entailment ~ (1|CorpusID) 
                        + cosine_similarity * tfidf_cosine_similarity
                        + tfidf_cosine_similarity * seg2_given_seg1_perplexity
                        + prag_distance 
                        + seg2_given_seg1_perplexity 
                        * seg1_perplexity 
                        + seg2_given_seg1_perplexity 
                        * seg2_perplexity
                        + seg2_n_tokens
                        + seg1_n_tokens * seg2_given_seg1_perplexity
                        + seg1_impScore
                        + seg2_impScore
                        #+ tfidf_cosine_similarity 
                        , data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))
```

```{r}
library(car)

anova(model_interact_tfidf_per_entailment, model_interact_seg1_len_per_entailment, test = "Chisq")

print("R2 with tfidf x perplexity distance interaction")
performance::r2_nakagawa(model_interact_tfidf_per_entailment)

print("R2 seg1_len x perplexity interaction")
performance::r2_nakagawa(model_interact_seg1_len_per_entailment)

vif(model_interact_tfidf_per_entailment)
vif(model_interact_seg1_len_per_entailment)
#vif(model_interact_with_length)
```

## Interact ImpScore x pragmatic distance

```{r}
model_interact_imp_entailment <- glmer(is_entailment ~ (1|CorpusID) 
                        + cosine_similarity * tfidf_cosine_similarity
                        + tfidf_cosine_similarity * seg2_given_seg1_perplexity
                        + prag_distance * seg1_impScore
                        + prag_distance * seg2_impScore
                        + seg2_given_seg1_perplexity * seg1_perplexity 
                        + seg2_given_seg1_perplexity * seg2_perplexity
                        + seg2_n_tokens
                        + seg1_n_tokens * seg2_given_seg1_perplexity
                        #+ seg1_impScore
                        #+ seg2_impScore
                        #+ tfidf_cosine_similarity 
                        , data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))
```

```{r}
library(car)

anova(model_interact_seg1_len_per_entailment, model_interact_imp_entailment, test = "Chisq")

print("R2 seg1_len x perplexity interaction")
performance::r2_nakagawa(model_interact_seg1_len_per_entailment)

print("R2 with imp x prag distance")
performance::r2_nakagawa(model_interact_imp_entailment)

vif(model_interact_seg1_len_per_entailment)
vif(model_interact_imp_entailment)
#vif(model_interact_with_length)
```

## Three-way interaction cos x tfidf x per

```{r}
model_interact_three_way_entailment <- glmer(is_entailment ~ (1|CorpusID) 
                        + cosine_similarity * tfidf_cosine_similarity * seg2_given_seg1_perplexity
                        + prag_distance * seg1_impScore
                        + prag_distance * seg2_impScore
                        + seg2_given_seg1_perplexity 
                        * seg1_perplexity 
                        + seg2_given_seg1_perplexity 
                        * seg2_perplexity
                        + seg2_n_tokens
                        + seg1_n_tokens * seg2_given_seg1_perplexity
                        #+ seg1_impScore
                        + seg2_impScore
                        #+ tfidf_cosine_similarity 
                        , data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))
```


```{r}
library(car)

anova(model_interact_imp_entailment, model_interact_three_way_entailment, test = "Chisq")

print("R2 with imp x prag distance")
performance::r2_nakagawa(model_interact_imp_entailment)

print("R2 3-way cos x tfidf x per interaction")
performance::r2_nakagawa(model_interact_three_way_entailment)

vif(model_interact_imp_entailment)
vif(model_interact_three_way_entailment)
#vif(model_interact_with_length)
```

## Interact 3-way tfidf x prag x per
```{r}
model_interact_three_way_prag_entailment <- glmer(is_entailment ~ (1|CorpusID) 
                        + cosine_similarity * tfidf_cosine_similarity * seg2_given_seg1_perplexity
                        + tfidf_cosine_similarity * seg2_given_seg1_perplexity * prag_distance
                        + prag_distance * seg1_impScore
                        + prag_distance * seg2_impScore
                        + seg2_given_seg1_perplexity 
                        * seg1_perplexity 
                        + seg2_given_seg1_perplexity 
                        * seg2_perplexity
                        + seg2_n_tokens
                        + seg1_n_tokens * seg2_given_seg1_perplexity
                        #+ seg1_impScore
                        #+ seg2_impScore
                        #+ tfidf_cosine_similarity 
                        , data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))
```

```{r}
library(car)

anova(model_interact_three_way_entailment, model_interact_three_way_prag_entailment, test = "Chisq")

print("R2 3-way cos x tfidf x per interaction")
performance::r2_nakagawa(model_interact_three_way_entailment)

print("R2 3-way tfidf x per x prag")
performance::r2_nakagawa(model_interact_three_way_prag_entailment)

vif(model_interact_three_way_entailment)
vif(model_interact_three_way_prag_entailment)
#vif(model_interact_with_length)
```

## Inspect current best fit -> three way cos x tfidf x per

We use an ANOVA to identify low-significance features
```{r}
Anova(model_interact_three_way_entailment)
```

```{r}
model_interact_three_way_refined_entailment <- glmer(is_entailment ~ (1|CorpusID) 
                        + cosine_similarity * tfidf_cosine_similarity * seg2_given_seg1_perplexity
                        #+ prag_distance * seg1_impScore
                        + prag_distance * seg2_impScore
                        + seg2_given_seg1_perplexity 
                        * seg1_perplexity 
                        + seg2_given_seg1_perplexity 
                        * seg2_perplexity
                        + seg2_n_tokens
                        + seg1_n_tokens * seg2_given_seg1_perplexity
                        #+ seg1_impScore
                        + seg2_impScore
                        #+ tfidf_cosine_similarity 
                        , data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))
```

```{r}
library(car)

anova(model_interact_three_way_entailment, model_interact_three_way_refined_entailment, test = "Chisq")

print("R2 3-way cos x tfidf x per interaction")
performance::r2_nakagawa(model_interact_three_way_entailment)

print("R2 3-way cos x tfidf x per interaction refined")
performance::r2_nakagawa(model_interact_three_way_refined_entailment)

vif(model_interact_three_way_entailment)
vif(model_interact_three_way_refined_entailment)
#vif(model_interact_with_length)
```


```{r}
Anova(model_interact_three_way_refined_entailment)
```
```{r}
library(broom.mixed)
library(dplyr)

effects_table <- tidy(model_interact_three_way_refined, effects = "fixed", conf.int = TRUE) %>%
  arrange(p.value) %>%
  mutate(significance = case_when(
    p.value < 0.001 ~ "***",
    p.value < 0.01  ~ "**",
    p.value < 0.05  ~ "*",
    TRUE ~ ""
  ))

effects_table
```

# Step 3: Check remaining feature blocks
```{r}
library(lme4)
library(performance)
library(broom.mixed)
library(purrr)

# Define your dropped blocks from earlier
blocks_dropped <- list(
  syntactic = c("seg1_dependency_distance_mean", "seg2_dependency_distance_mean",
                "seg1_prop_adjacent_dependency_relation_mean", "seg2_prop_adjacent_dependency_relation_mean",
                "seg1_avg_subclauses", "seg2_avg_subclauses",
                "seg1_avg_tree_depth", "seg2_avg_tree_depth"),
  structure = c("seg1_sentence_length_mean", "seg2_sentence_length_mean"),
  sentiment = c("seg1_senti", "seg2_senti"),
  lexical_markers = c("seg1_has_deduction", "seg2_has_deduction",
                      "seg1_has_refutation", "seg2_has_refutation", 
                      "seg1_has_condition", "seg2_has_condition",
                      "seg1_has_explanation", "seg2_has_explanation", 
                      "seg1_has_adverb_of_emphasis", "seg2_has_adverb_of_emphasis",
                      "seg1_has_justification", "seg2_has_justification", 
                      "seg1_has_modal_verbs", "seg2_has_modal_verbs")
)

# Loop through them and test one-at-a-time addition
block_refit_results_entailment <- map_df(names(blocks_dropped), function(b) {
  predictors <- paste(blocks_dropped[[b]], collapse = " + ")
  m_block <- update(model_interact_three_way_refined, paste(". ~ . +", predictors))
  comp <- anova(model_interact_three_way_refined, m_block, test = "Chisq")
  tibble(
    block = b,
    delta_AIC = AIC(model_interact_three_way_refined) - AIC(m_block),
    LRT_p = comp$`Pr(>Chisq)`[2],
    R2_marg_gain = performance::r2_nakagawa(m_block)$R2_marginal -
                   performance::r2_nakagawa(model_interact_three_way_refined)$R2_marginal
  )
})

block_refit_results
```

```{r}
#summary(update(model_interact_three_way_refined, . ~ . + seg1_avg_tree_depth + seg2_avg_tree_depth))
car::Anova(update(model_interact_three_way_refined_entailment, . ~ . #+ seg1_has_deduction
                                                          + seg2_has_deduction
                                                          #+ seg1_has_refutation
                                                          + seg2_has_refutation
                                                          + seg1_has_condition
                                                          #+ seg2_has_condition
                                                          #+ seg1_has_explanation
                                                          + seg2_has_explanation
                                                          #+ seg1_has_adverb_of_emphasis
                                                          + seg2_has_adverb_of_emphasis
                                                          #+ seg1_has_justification
                                                          + seg2_has_justification
                                                          + seg1_has_modal_verbs
                                                          + seg2_has_modal_verbs), type="III")
```

```{r}
car::Anova(update(model_interact_three_way_refined_entailment, . ~ . + seg1_sentence_length_mean
                                                          + seg2_sentence_length_mean), type="III")
```

```{r}
car::Anova(update(model_interact_three_way_refined_entailment, . ~ . + seg1_senti + seg2_senti), type="III")
```

```{r}
car::Anova(update(model_interact_three_way_refined_entailment, . ~ . + seg1_dependency_distance_mean
                                                          + seg2_dependency_distance_mean
                                                          + seg1_prop_adjacent_dependency_relation_mean
                                                          + seg2_prop_adjacent_dependency_relation_mean
                                                          + seg1_avg_subclauses
                                                          + seg2_avg_subclauses
                                                          + seg1_avg_tree_depth
                                                          + seg2_avg_tree_depth), type="III")
```
## Add sinificant features from other blocks
```{r}
model_final_expanded_interactions_entailment <- glmer(
  is_entailment ~ (1 | CorpusID) 
  + cosine_similarity * tfidf_cosine_similarity 
  + seg2_given_seg1_perplexity * cosine_similarity 
  + prag_distance * seg2_impScore 
  + seg2_given_seg1_perplexity * seg1_perplexity 
  + seg2_given_seg1_perplexity * seg2_perplexity 
  + seg2_n_tokens * seg2_sentence_length_mean 
  + seg2_n_tokens * seg2_avg_tree_depth 
  + seg2_sentence_length_mean * seg2_avg_tree_depth 
  + seg1_sentence_length_mean * seg1_avg_tree_depth
  + seg1_senti 
  + seg2_senti 
  + seg2_has_refutation 
  + seg1_has_condition
  + seg1_has_modal_verbs
  + seg2_has_modal_verbs
  , data = hard_threshold_complete,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa")
)

```

```{r}
anova(model_interact_three_way_refined_entailment, model_final_expanded_interactions_entailment, test = "Chisq")
print("R2 for model_interact_three_way_refined")
performance::r2_nakagawa(model_interact_three_way_refined_entailment)
print("R2 for model_final_expanded_interactions")
performance::r2_nakagawa(model_final_expanded_interactions_entailment)
```


## Interaction tree depth
```{r}
model_final_expanded_tree_depth_entailment <- glmer(
  is_entailment ~ (1 | CorpusID) 
  + cosine_similarity * tfidf_cosine_similarity 
  + seg2_given_seg1_perplexity * cosine_similarity 
  + prag_distance * seg2_impScore 
  + seg2_given_seg1_perplexity * seg1_perplexity 
  + seg2_given_seg1_perplexity * seg2_perplexity 
  # + seg2_n_tokens * seg2_sentence_length_mean 
  #+ seg2_n_tokens * seg2_avg_tree_depth 
  #+ seg2_sentence_length_mean * 
  + seg2_avg_tree_depth 
  + seg1_avg_tree_depth
  #+ seg1_sentence_length_mean 
  + seg1_senti 
  + seg2_senti 
  + seg2_has_refutation 
  + seg1_has_condition
  + seg1_has_modal_verbs
  + seg2_has_modal_verbs
  , data = hard_threshold_complete,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa")
)

```

```{r}
anova(model_final_expanded_interactions_entailment, model_final_expanded_tree_depth_entailment, test = "Chisq")
message("R2 for model_final_expanded_interactions")
performance::r2_nakagawa(model_final_expanded_interactions_entailment)
message("R2 for model_final_expanded_tree_depth")
performance::r2_nakagawa(model_final_expanded_tree_depth_entailment)
```

## sentence lenght vs tokens as length measure
```{r}
model_final_expanded_sentence_length_entailment <- glmer(
  is_entailment ~ (1 | CorpusID) 
  + cosine_similarity * tfidf_cosine_similarity 
  + seg2_given_seg1_perplexity * cosine_similarity 
  + prag_distance * seg2_impScore 
  + seg2_given_seg1_perplexity * seg1_perplexity 
  + seg2_given_seg1_perplexity * seg2_perplexity 
  # + seg2_n_tokens * seg2_sentence_length_mean 
  #+ seg2_n_tokens * seg2_avg_tree_depth 
  + seg2_sentence_length_mean #* 
  #+ seg2_avg_tree_depth 
  + seg1_sentence_length_mean 
  + seg1_senti 
  + seg2_senti 
  + seg2_has_refutation 
  + seg1_has_condition
  + seg1_has_modal_verbs
  + seg2_has_modal_verbs
  , data = hard_threshold_complete,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa")
)

```

```{r}
model_final_expanded_n_tokens_entailment <- glmer(
  is_entailment ~ (1 | CorpusID) 
  + cosine_similarity * tfidf_cosine_similarity 
  + seg2_given_seg1_perplexity * cosine_similarity 
  + prag_distance * seg2_impScore 
  + seg2_given_seg1_perplexity * seg1_perplexity 
  + seg2_given_seg1_perplexity * seg2_perplexity 
  + seg2_n_tokens 
  #* seg2_sentence_length_mean 
  #+ seg2_n_tokens * seg2_avg_tree_depth 
  #+ seg2_sentence_length_mean #* 
  #+ seg2_avg_tree_depth 
  #+ seg1_sentence_length_mean 
  + seg1_senti 
  + seg2_senti 
  + seg2_has_refutation 
  + seg1_has_condition
  + seg1_has_modal_verbs
  + seg2_has_modal_verbs
  , data = hard_threshold_complete,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa")
)

```

```{r}
anova(model_final_expanded_n_tokens_entailment, model_final_expanded_sentence_length_entailment, test = "Chisq")
message("R2 for model_final_expanded_sentence_length")
performance::r2_nakagawa(model_final_expanded_sentence_length_entailment)
message("R2 for model_final_expanded_n_tokens")
performance::r2_nakagawa(model_final_expanded_n_tokens_entailment)
```

## 3-way interaction

```{r}
model_final_expanded_n_tokens_threeWay_entailment <- glmer(
  is_entailment ~ (1 | CorpusID) 
  + cosine_similarity * tfidf_cosine_similarity 
  * seg2_given_seg1_perplexity #* cosine_similarity 
  + prag_distance * seg2_impScore 
  + seg2_given_seg1_perplexity * seg1_perplexity 
  + seg2_given_seg1_perplexity * seg2_perplexity 
  + seg2_n_tokens #* seg2_sentence_length_mean 
  #+ seg2_n_tokens * seg2_avg_tree_depth 
  #+ seg2_sentence_length_mean * seg2_avg_tree_depth 
  #+ seg1_sentence_length_mean * seg1_avg_tree_depth
  + seg1_senti 
  + seg2_senti 
  + seg2_has_refutation 
  + seg1_has_condition
  + seg1_has_modal_verbs
  + seg2_has_modal_verbs
  , data = hard_threshold_complete,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa")
)
```

```{r}
anova(model_final_expanded_n_tokens_entailment, model_final_expanded_n_tokens_threeWay_entailment, test = "Chisq")

message("R2 for model_final_expanded_n_tokens")
performance::r2_nakagawa(model_final_expanded_n_tokens_entailment)
message("R2 for model_final_expanded_n_tokens_threeWay")
performance::r2_nakagawa(model_final_expanded_n_tokens_threeWay_entailment)
```


```{r}
model_final_expanded_n_tokens_noMarginal_entailment <- glmer(
  is_entailment ~ (1 | CorpusID) 
  + cosine_similarity * tfidf_cosine_similarity 
  #* seg2_given_seg1_perplexity #* cosine_similarity 
  + prag_distance * seg2_impScore 
  + seg2_given_seg1_perplexity * seg1_perplexity 
  + seg2_given_seg1_perplexity * seg2_perplexity 
  + seg2_n_tokens #* seg2_sentence_length_mean 
  #+ seg2_n_tokens * seg2_avg_tree_depth 
  #+ seg2_sentence_length_mean * seg2_avg_tree_depth 
  #+ seg1_sentence_length_mean * seg1_avg_tree_depth
  + seg1_senti 
  + seg2_senti 
  + seg2_has_refutation 
  + seg1_has_condition
  + seg1_has_modal_verbs
  + seg2_has_modal_verbs
  , data = hard_threshold_complete,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa")
)
```

```{r}
anova(model_final_expanded_n_tokens_threeWay_entailment, model_final_expanded_n_tokens_noMarginal_entailment, test = "Chisq")

message("R2 for model_final_expanded_n_tokens_threeWay")
performance::r2_nakagawa(model_final_expanded_n_tokens_threeWay_entailment)
message("R2 for model_final_expanded_n_tokens_noMarginal")
performance::r2_nakagawa(model_final_expanded_n_tokens_noMarginal_entailment)
```


```{r}
car::Anova(model_final_expanded_interactions_entailment, type="III")
car::Anova(model_final_expanded_n_tokens_threeWay_entailment, type="III")
car::Anova(model_final_expanded_n_tokens_noMarginal_entailment, type="III")
```

```{r}
performance::compare_performance(model_final_expanded_interactions_entailment, model_final_expanded_n_tokens_threeWay_entailment, model_final_expanded_n_tokens_noMarginal_entailment)
```

```{r}
performance::check_collinearity(model_final_expanded_interactions_entailment)
performance::check_collinearity(model_final_expanded_n_tokens_threeWay_entailment)
performance::check_collinearity(model_final_expanded_n_tokens_noMarginal_entailment)
```
We choose the model "model_final_expanded_n_tokens_threeWay_entailment" as the final model. The "model_final_expanded_interactions_entailment" might have the higher R2, but this model shows strong collinearity, which makes the chosen final model the better balance between performance and stability.


```{r}
model_final_entailment <- model_final_expanded_n_tokens_threeWay_entailment
# --- This is your MAIN results table ---
# It's simple, clean, and you already have it.
library(car)
model_significance <- car::Anova(model_final_entailment, type = "III")
print(model_significance)

# --- This table goes in your APPENDIX ---
library(broom.mixed)
appendix_table <- tidy(model_final_entailment,
                       effects = "fixed",
                       exponentiate = TRUE,
                       conf.int = TRUE)

print(appendix_table)

# --- This is the ONE-LINE solution ---
library(sjPlot)

# This command creates a beautiful, publication-ready
# HTML table right in your RStudio viewer.
tab_model(
  model_final_entailment, 
  dv.labels = "Predicting Entailment (is_entailment = 1)",
  p.style = "scientific"
)
```

# Extra stuff



```{r}
model_final_expanded_scale_len <- glmer(
  is_entailment ~ (1 | CorpusID) 
  + cosine_similarity * tfidf_cosine_similarity 
  + seg2_given_seg1_perplexity * cosine_similarity 
  + prag_distance * seg2_impScore 
  + seg2_given_seg1_perplexity * seg1_perplexity 
  + seg2_given_seg1_perplexity * seg2_perplexity 
  + scale(seg2_n_tokens + seg2_sentence_length_mean + seg2_avg_tree_depth)
  + seg1_senti + seg2_senti 
  + seg2_has_refutation 
  + seg1_has_condition 
  + seg1_has_modal_verbs 
  + seg2_has_modal_verbs
  ,data = hard_threshold_complete,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa")
)
```

```{r}
anova(model_interact_three_way_refined, model_final_expanded, model_final_expanded_scale_len, test = "Chisq")
performance::r2_nakagawa(model_interact_three_way_refined)
performance::r2_nakagawa(model_final_expanded)
performance::r2_nakagawa(model_final_expanded_scale_len)
```
```{r}
car::vif(model_final_expanded_scale_len)
```
```{r}
library(sjPlot)

library(parameters)
params <- model_parameters(
  model_final_entailment,
  ci = 0.95,
  standardize = "refit",
  bootstrap = FALSE
)
print(params, pretty_names = TRUE, stars = TRUE)

# Example with your best model
tab_model(
  ,
  show.ci = TRUE,
  show.std = TRUE,
  show.se = TRUE,
  show.stat = TRUE,
  show.re.var = TRUE,    # shows random effects
  p.style = "stars",      # adds *, **, *** significance
  digits = 3,
  ci.lvl = 0.95,
  title = "Mixed-Effects Logistic Regression Predicting Support Relations",
  dv.labels = "is_support",
  pred.labels = c(
    "Seg2 has deduction", "Seg1 has refutation", "Seg2 has refutation",
    "Seg1 has condition", "Seg2 has adverb of emphasis", "Seg1 has justification",
    "Seg2 has justification", "Seg2 sentiment", "Seg2 avg. subclauses",
    "Seg2 dependency distance mean", "Seg2 prop. adjacent dependency",
    "Seg2 avg. tree depth", "Seg1 avg. tree depth",
    "Seg1 dependency distance mean", "Seg1 perplexity",
    "Seg2→Seg1 perplexity", "TF-IDF cosine similarity",
    "Seg1 implicature score", "Seg2 implicature score",
    "Pragmatic distance", "Cosine similarity",
    "Seg2 avg. subclauses × Seg2 dependency distance mean",
    "Seg2 avg. subclauses × Seg2 avg. tree depth",
    "Seg1 avg. tree depth × Seg1 dependency distance mean",
    "Pragmatic distance × Cosine similarity"
  )
)
```


```{r}
car::Anova(model_final_expanded_scale_len, type="III")
```
```{r}
performance::check_collinearity(model_final_expanded_scale_len)
```

