---
title: "support_stat_modelFitting"
author: "Henrike Beyer"
date: "2025-11-07"
output: html_document
---

```{r}
library(lme4)
library(broom.mixed)
library(dplyr)
```

# Data Loading and Preparation

In this section, we load in the data, transform all relevant character vectors to factors, and clean up NAs in the data.

```{r read Data}
hard_threshold <- read.csv("D:\\OneDrive - University of Dundee\\PhD\\ARG-NLI_project\\feature_analysis\\automatic\\new_features\\corpus_threshold_hard_0.999_standardised_and_log_scaled.csv")

```

```{r transform character vectors to factors}
factor_vars <- c(
  "seg1_has_deduction", "seg2_has_deduction",
  "seg1_has_refutation", "seg2_has_refutation", 
  "seg1_has_condition", "seg2_has_condition",
  "seg1_has_explanation", "seg2_has_explanation", 
  "seg1_has_adverb_of_emphasis", "seg2_has_adverb_of_emphasis",
  "seg1_has_justification", "seg2_has_justification", 
  "seg1_has_modal_verbs", "seg2_has_modal_verbs",
  "seg1_senti", "seg2_senti", "CorpusID"
)

hard_threshold[factor_vars] <- lapply(hard_threshold[factor_vars], function(x) as.factor(as.character(x)))
```


```{r define outcomes}
is_entailment <- "is_entailment"
is_contradiction <- "is_contradiction"
is_neutral <- "is_neutral"
is_support <- "is_support"
is_attack <- "is_attack"
support_is_entailment <- "arg_nli_support"
attack_is_contradiction <- "arg_nli_attack"

outcome <- is_support
```


```{r clean up NAs}
# Drop rows with any missing values
hard_threshold_complete <- hard_threshold[complete.cases(hard_threshold), ]
```

# Step 1: Block model fitting
In the first step, we split up the entire set of features into 8 linguistically motivated blocks and fit one model for each block that includes these features plus the number of tokens as a baseline feature. The gain in AIC is then compared for each block with the baseline model with the number of tokens as the only feature. Since we cannot guarantee that there is no interdependence introduced by the different corpora, the corpus is always used as a random effect.

```{r define blocks}
blocks <- list(
  lexical_markers = c("seg1_has_deduction", "seg2_has_deduction",
                      "seg1_has_refutation", "seg2_has_refutation", 
                      "seg1_has_condition", "seg2_has_condition",
                      "seg1_has_explanation", "seg2_has_explanation", 
                      "seg1_has_adverb_of_emphasis", "seg2_has_adverb_of_emphasis",
                      "seg1_has_justification", "seg2_has_justification", 
                      "seg1_has_modal_verbs", "seg2_has_modal_verbs"),
  lexical_sim = c("tfidf_cosine_similarity"),
  semantic = c("cosine_similarity"),
  structure = c("seg1_sentence_length_mean", "seg2_sentence_length_mean"),
  #                     "seg1_n_tokens", "seg2_n_tokens"),
  syntactic = c("seg1_dependency_distance_mean", "seg2_dependency_distance_mean",
                "seg1_prop_adjacent_dependency_relation_mean", "seg2_prop_adjacent_dependency_relation_mean",
                "seg1_avg_subclauses", "seg2_avg_subclauses",
                "seg1_avg_tree_depth", "seg2_avg_tree_depth"),
  perplexity = c("seg1_perplexity", "seg2_perplexity", #"seg1_given_seg2_perplexity", 
                 "seg2_given_seg1_perplexity"),
  indirectness = c("seg1_impScore", "seg2_impScore", "prag_distance"),
  sentiment = c("seg1_senti", "seg2_senti")
)

interactions <- list(
  c("seg1_avg_subclauses", "seg1_sentence_length_mean"),
  c("seg2_avg_subclauses", "seg2_sentence_length_mean"),
  c("seg1_avg_tree_depth", "seg1_avg_subclauses"),
  c("seg2_avg_tree_depth", "seg2_avg_subclauses"),
  c("seg1_avg_tree_depth", "seg1_sentence_length_mean"),
  c("seg2_avg_tree_depth", "seg2_sentence_length_mean"),
  c("seg1_dependency_distance_mean", "seg1_sentence_length_mean"),
  c("seg2_sentence_length_mean", "seg2_dependency_distance_mean"),
  c("seg1_dependency_distance_mean", "seg1_avg_subclauses"),
  c("prag_distance", "cosine_similarity"),
  c("tfidf_cosine_similarity", "cosine_similarity"),
  c("seg2_dependency_distance_mean", "seg2_avg_subclauses"),
  c("seg2_sentence_length_mean", "seg2_n_tokens"),
  c("seg2_avg_subclauses", "seg2_n_tokens"),
  c("seg1_avg_subclauses", "seg1_n_tokens"),
  c("seg1_sentence_length_mean", "seg1_n_tokens"),
  c("seg2_dependency_distance_mean", "seg2_n_tokens"),
  c("tfidf_cosine_similarity", "prag_distance"),
  c("seg1_avg_tree_depth", "seg1_dependency_distance_mean"),
  c("seg2_avg_tree_depth", "seg2_n_tokens"),
  c("seg1_dependency_distance_mean", "seg1_n_tokens"),
  c("seg1_avg_tree_depth", "seg1_n_tokens"),
  c("seg2_dependency_distance_mean", "seg2_prop_adjacent_dependency_relation_mean"),
  c("seg1_prop_adjacent_dependency_relation_mean", "seg1_dependency_distance_mean")
)

```

```{r helper variables to formula}
vars_to_formula <- function(vars, interactions = NULL) {
  f <- paste(vars, collapse = " + ")
  if (!is.null(interactions) && length (interactions) > 0) {
    int_terms <- sapply(interactions, function(x) paste(x, collapse = ":"))
  }
  f
}
```

```{r}
fit_block_models_safe <- function(df, outcome, blocks, random = "(1 | CorpusID)",
                                  base_controls = c("seg1_n_tokens", "seg2_n_tokens"),
                                  family = binomial) {
  library(lme4)
  library(performance)
  library(dplyr)
  
  models <- list()
  results <- tibble()
  
  # baseline model
  base_formula <- paste0(outcome, " ~ ", paste(base_controls, collapse = " + "), " + ", random)
  message("Fitting baseline model: ", base_formula)
  m_base <- glmer(as.formula(base_formula), data = df,
                  family = family, control = glmerControl(optimizer = "bobyqa"))
  models[["baseline"]] <- m_base
  results <- bind_rows(results, tibble(
    block = "baseline", AIC = AIC(m_base),
    delta_AIC = NA, LRT = NA, p = NA, r2_marg = NA
  ))
  
  # loop over blocks
  for (b in names(blocks)) {
    vars <- c(base_controls, blocks[[b]])
    fstring <- paste0(outcome, " ~ ", paste(vars, collapse = " + "), " + ", random)
    message("Fitting block-only model: ", b)
    
    m <- tryCatch(
      glmer(as.formula(fstring), data = df, family = family,
            control = glmerControl(optimizer = "bobyqa")),
      error = function(e) { warning("Model failed for block ", b, ": ", e$message); return(NULL) }
    )
    
    if (!is.null(m)) {
      models[[b]] <- m
      comp <- tryCatch(anova(m_base, m, test = "Chisq"), error = function(e) NULL)
      r2 <- tryCatch(performance::r2_nakagawa(m)$R2_marginal, error = function(e) NA)
      
      row <- tibble(
        block = b,
        AIC = AIC(m),
        delta_AIC = if (!is.null(comp)) AIC(m_base) - AIC(m) else NA,
        LRT = if (!is.null(comp)) comp$Chisq[2] else NA,
        p = if (!is.null(comp)) comp$`Pr(>Chisq)`[2] else NA,
        r2_marg = r2
      )
      results <- bind_rows(results, row)
    } else {
      results <- bind_rows(results, tibble(
        block = b, AIC = NA, delta_AIC = NA, LRT = NA, p = NA, r2_marg = NA
      ))
    }
  }
  
  list(models = models, summary = results)
}
```


```{r}
block_results_support <- fit_block_models_safe(hard_threshold_complete, outcome = outcome, blocks = blocks)

block_results_support$summary %>%
  arrange(AIC)  # smaller AIC = better fit
```

```{r}
library(ggplot2)
ggplot(block_results_support$summary[-1,], aes(x = reorder(block, -delta_AIC), y = delta_AIC)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(x = "Feature Block", y = "ΔAIC improvement over baseline",
       title = "Relative contribution of feature blocks")
```
```{r}
extract_block_details <- function(block_results_support) {
  library(broom.mixed)
  lapply(names(block_results_support$models), function(b) {
    m <- block_results_support$models[[b]]
    if (is.null(m)) return(NULL)
    broom.mixed::tidy(m, effects = "fixed", conf.int = TRUE, exponentiate = TRUE) %>%
      mutate(block = b)
  }) %>% bind_rows() %>% filter(term != "(Intercept)")
}
```

```{r}
block_details_support <- extract_block_details(block_results_support)
```

```{r}
library(car)
car::Anova(block_results_support$models[["lexical_markers"]])
car::Anova(block_results_support$models[["sentiment"]])
car::Anova(block_results_support$models[["syntactic"]])
car::Anova(block_results_support$models[["lexical_sim"]])
```


```{r}
strong_blocks_support <- c("lexical_markers", "sentiment", "syntactic", "lexical_sim")

sig_features_support <- block_details_support %>%
  filter(block %in% strong_blocks_support,
         p.value < 0.05,
         abs(estimate - 1) > 0.05) %>%
  arrange(block, p.value) %>%
  select(block, term, estimate, p.value, conf.low, conf.high)

sig_features_support
```

```{r}
unique(sig_features_support$term)
```


```{r}
final_features_support <- c("seg2_n_tokens",
                    "seg2_has_deduction",
                    "seg1_has_refutation",
                    "seg2_has_refutation",
                    "seg1_has_condition",
                    "seg2_has_condition",
                    "seg2_has_adverb_of_emphasis",
                    "seg1_has_justification",
                    "seg2_has_justification",
                    "seg2_has_modal_verbs",
                    "seg2_senti",
                    "seg1_n_tokens",
                    "seg1_dependency_distance_mean",
                    "seg2_dependency_distance_mean",
                    "seg1_prop_adjacent_dependency_relation_mean",
                    "seg2_prop_adjacent_dependency_relation_mean",
                    "seg2_avg_subclauses",
                    "seg1_avg_tree_depth",
                    "seg2_avg_tree_depth",
                    "tfidf_cosine_similarity"
)

f_final_support <- as.formula(
  paste("is_support ~", paste(final_features_support, collapse = " + "), "+ (1 | CorpusID)")
)

model_baseline_support <-glmer(is_support ~ (1|CorpusID) + seg1_n_tokens + seg2_n_tokens, data = hard_threshold_complete,
                       family = binomial,
                       control = glmerControl(optimizer = "bobyqa"))

model_block_support <- glmer(f_final, data = hard_threshold_complete,
                     family = binomial,
                     control = glmerControl(optimizer = "bobyqa"))
```

```{r}
model_without_length_support <- glmer(is_support ~ (1|CorpusID) 
                        #+ seg2_n_tokens
                        + seg2_has_deduction
                        + seg1_has_refutation
                        + seg2_has_refutation
                        + seg1_has_condition
                        + seg2_has_condition
                        + seg2_has_adverb_of_emphasis
                        + seg1_has_justification
                        + seg2_has_justification
                        + seg2_has_modal_verbs
                        + seg2_senti
                        #+ seg1_n_tokens
                        + seg1_dependency_distance_mean
                        + seg2_dependency_distance_mean
                        + seg1_prop_adjacent_dependency_relation_mean
                        + seg2_prop_adjacent_dependency_relation_mean
                        + seg2_avg_subclauses
                        + seg1_avg_tree_depth
                        + seg2_avg_tree_depth
                        + tfidf_cosine_similarity
                        , data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))
```

```{r}
library(car)

anova(model_baseline_support, model_blocks_support, model_without_length_support, test = "Chisq")

print("R2 model blocks")
performance::r2_nakagawa(model_blocks_support)

print("R2 without lenght")
performance::r2_nakagawa(model_without_length_support)

```

```{r}
# --- 1. Load the necessary libraries ---
library(tidyverse)
library(fastDummies) # For handling multi-level factors

# --- 2. Get predictor names ---
# Use the formula for your 22-predictor 'model_final' for 'is_support'
fe_formula <- formula(model_without_length_support, fixed.only = TRUE)
raw_predictors <- all.vars(fe_formula)[-1] 

# --- 3. Create numeric-only dataframe for correlation ---
model_data_numeric <- hard_threshold_complete %>%
  
  # Select only the predictors that are actually in the final model
  select(all_of(raw_predictors)) %>%
  
  # *** NEW DUMMY VARIABLE STEP ***
  # This will find any factor/character columns (like 'seg2_senti')
  # and create new 0/1 columns for each level (e.g., seg2_senti_POSITIVE)
  dummy_cols(remove_selected_columns = TRUE, # Removes the original factor
             remove_first_dummy = FALSE)     # Keep all levels for the matrix

# --- 4. Calculate the correlation matrix ---
predictor_cor_matrix <- cor(model_data_numeric, use = "pairwise.complete.obs")

# --- 5. Filter for strong correlations ---
# (The rest of your code from before is perfect)
strong_correlations <- as.data.frame(predictor_cor_matrix) %>%
  rownames_to_column("Var1") %>%
  pivot_longer(
    cols = -Var1,
    names_to = "Var2",
    values_to = "Correlation"
  ) %>%
  filter(abs(Correlation) > 0.4 & abs(Correlation) < 1.0 & Var1 < Var2) %>%
  arrange(desc(abs(Correlation)))

# --- 6. Print the final, clean table ---
print(strong_correlations)
```

```{r}
model_interact_subclauses_tree_support <-  glmer(is_support ~ (1|CorpusID) 
                        #+ seg2_n_tokens
                        + seg2_has_deduction
                        + seg1_has_refutation
                        + seg2_has_refutation
                        + seg1_has_condition
                        + seg2_has_condition
                        + seg2_has_adverb_of_emphasis
                        + seg1_has_justification
                        + seg2_has_justification
                        + seg2_has_modal_verbs
                        + seg2_senti
                        #+ seg1_n_tokens
                        + seg1_dependency_distance_mean
                        + seg2_dependency_distance_mean
                        + seg1_prop_adjacent_dependency_relation_mean
                        + seg2_prop_adjacent_dependency_relation_mean
                        + seg2_avg_subclauses * seg2_avg_tree_depth
                        + seg1_avg_tree_depth
                        + tfidf_cosine_similarity
                        , data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))
```

```{r}
library(car)

anova(model_without_length_support, model_interact_subclauses_tree_support, test = "Chisq")

print("R2 without lenght")
performance::r2_nakagawa(model_without_length_support)

print("R2 model subclauses x tree depth")
performance::r2_nakagawa(model_interact_subclauses_tree_support)
```

```{r}
model_interact_subclauses_dep_support <-  glmer(is_support ~ (1|CorpusID) 
                        #+ seg2_n_tokens
                        + seg2_has_deduction
                        + seg1_has_refutation
                        + seg2_has_refutation
                        + seg1_has_condition
                        + seg2_has_condition
                        + seg2_has_adverb_of_emphasis
                        + seg1_has_justification
                        + seg2_has_justification
                        + seg2_has_modal_verbs
                        + seg2_senti
                        #+ seg1_n_tokens
                        + seg1_dependency_distance_mean
                        #+ seg2_dependency_distance_mean
                        + seg2_avg_subclauses *	seg2_dependency_distance_mean
                        + seg1_prop_adjacent_dependency_relation_mean
                        + seg2_prop_adjacent_dependency_relation_mean
                        + seg2_avg_subclauses * seg2_avg_tree_depth
                        + seg1_avg_tree_depth
                        + tfidf_cosine_similarity
                        , data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))
```

```{r}
library(car)

anova(model_interact_subclauses_tree_support, model_interact_subclauses_dep_support, test = "Chisq")

print("R2 model subclauses x tree depth")
performance::r2_nakagawa(model_interact_subclauses_tree_support)

print("R2 subclauses x dependencies")
performance::r2_nakagawa(model_interact_subclauses_dep_support)
```

```{r}
model_interact_seg1_tree_dep_support <-  glmer(is_support ~ (1|CorpusID) 
                        #+ seg2_n_tokens
                        + seg2_has_deduction
                        + seg1_has_refutation
                        + seg2_has_refutation
                        + seg1_has_condition
                        + seg2_has_condition
                        + seg2_has_adverb_of_emphasis
                        + seg1_has_justification
                        + seg2_has_justification
                        + seg2_has_modal_verbs
                        + seg2_senti
                        #+ seg1_n_tokens
                        #+ seg1_dependency_distance_mean
                        #+ seg2_dependency_distance_mean
                        + seg2_avg_subclauses *	seg2_dependency_distance_mean
                        + seg1_prop_adjacent_dependency_relation_mean
                        + seg2_prop_adjacent_dependency_relation_mean
                        + seg2_avg_subclauses * seg2_avg_tree_depth
                        + seg1_avg_tree_depth *	seg1_dependency_distance_mean
                        #+ seg1_avg_tree_depth
                        + tfidf_cosine_similarity
                        , data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))
```

```{r}
library(car)

anova(model_interact_subclauses_dep_support, model_interact_seg1_tree_dep_support, test = "Chisq")

print("R2 subclauses x dependencies")
performance::r2_nakagawa(model_interact_subclauses_dep_support)

print("R2 model dependency x tree depth")
performance::r2_nakagawa(model_interact_seg1_tree_dep_support)
```

```{r}
model_interact_seg2_tree_dep_support <-  glmer(is_support ~ (1|CorpusID) 
                        #+ seg2_n_tokens
                        + seg2_has_deduction
                        + seg1_has_refutation
                        + seg2_has_refutation
                        + seg1_has_condition
                        + seg2_has_condition
                        + seg2_has_adverb_of_emphasis
                        + seg1_has_justification
                        + seg2_has_justification
                        + seg2_has_modal_verbs
                        + seg2_senti
                        #+ seg1_n_tokens
                        #+ seg1_dependency_distance_mean
                        #+ seg2_dependency_distance_mean
                        + seg2_avg_subclauses *	seg2_dependency_distance_mean
                        + seg1_prop_adjacent_dependency_relation_mean
                        + seg2_prop_adjacent_dependency_relation_mean
                        + seg2_avg_subclauses * seg2_avg_tree_depth
                        + seg1_avg_tree_depth *	seg1_dependency_distance_mean
                        + seg2_avg_tree_depth *	seg2_dependency_distance_mean
                        #+ seg1_avg_tree_depth
                        + tfidf_cosine_similarity
                        , data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))
```

```{r}
library(car)

anova(model_interact_seg1_tree_dep_support, model_interact_seg2_tree_dep_support, test = "Chisq")

print("R2 model dependency x tree depth seg1")
performance::r2_nakagawa(model_interact_seg1_tree_dep_support)

print("R2 model dependency x tree depth seg2")
performance::r2_nakagawa(model_interact_seg2_tree_dep_support)
```

```{r}
model_interact_three_way_support <-  glmer(is_support ~ (1|CorpusID) 
                        #+ seg2_n_tokens
                        + seg2_has_deduction
                        + seg1_has_refutation
                        + seg2_has_refutation
                        + seg1_has_condition
                        + seg2_has_condition
                        + seg2_has_adverb_of_emphasis
                        + seg1_has_justification
                        + seg2_has_justification
                        + seg2_has_modal_verbs
                        + seg2_senti
                        #+ seg1_n_tokens
                        #+ seg1_dependency_distance_mean
                        #+ seg2_dependency_distance_mean
                        + seg2_avg_subclauses *	seg2_dependency_distance_mean * seg2_avg_tree_depth
                        + seg1_prop_adjacent_dependency_relation_mean
                        + seg2_prop_adjacent_dependency_relation_mean
                        #+ seg2_avg_subclauses * seg2_avg_tree_depth
                        + seg1_avg_tree_depth *	seg1_dependency_distance_mean
                        #+ seg2_avg_tree_depth *	seg2_dependency_distance_mean
                        #+ seg1_avg_tree_depth
                        + tfidf_cosine_similarity
                        , data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))
```

```{r}
library(car)

anova(model_interact_seg1_tree_dep_support, model_interact_three_way_support, test = "Chisq")

print("R2 model dependency x tree depth seg1")
performance::r2_nakagawa(model_interact_seg1_tree_dep_support)

print("R2 seg2 subclauses x dep x tree depth")
performance::r2_nakagawa(model_interact_three_way_support)
```


```{r}
Anova(model_interact_seg1_tree_dep_support)
Anova(model_interact_subclauses_dep_support)
```

```{r}
model_interact_seg1_tree_dep_refined_support <-  glmer(is_support ~ (1|CorpusID) 
                        #+ seg2_n_tokens
                        + seg2_has_deduction
                        + seg1_has_refutation
                        + seg2_has_refutation
                        + seg1_has_condition
                        #+ seg2_has_condition
                        + seg2_has_adverb_of_emphasis
                        + seg1_has_justification
                        + seg2_has_justification
                        #+ seg2_has_modal_verbs
                        + seg2_senti
                        #+ seg1_n_tokens
                        #+ seg1_dependency_distance_mean
                        #+ seg2_dependency_distance_mean
                        + seg2_avg_subclauses *	seg2_dependency_distance_mean
                        + seg1_prop_adjacent_dependency_relation_mean
                        + seg2_prop_adjacent_dependency_relation_mean
                        + seg2_avg_subclauses * seg2_avg_tree_depth
                        + seg1_avg_tree_depth *	seg1_dependency_distance_mean
                        #+ seg1_avg_tree_depth
                        + tfidf_cosine_similarity
                        , data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))
```

```{r}
library(car)

anova(model_interact_three_way_support, model_interact_seg1_tree_dep_refined_support, test = "Chisq")

print("R2 model dependency x tree depth seg1")
performance::r2_nakagawa(model_interact_seg1_tree_dep_support)

print("R2 model dependency x tree depth seg1 refined")
performance::r2_nakagawa(model_interact_seg1_tree_dep_refined_support)
```

```{r}
Anova(model_interact_seg1_tree_dep_refined_support)
```
```{r}
library(broom.mixed)
library(dplyr)

effects_table <- tidy(model_interact_seg1_tree_dep_refined_support, effects = "fixed", conf.int = TRUE) %>%
  arrange(p.value) %>%
  mutate(significance = case_when(
    p.value < 0.001 ~ "***",
    p.value < 0.01  ~ "**",
    p.value < 0.05  ~ "*",
    TRUE ~ ""
  ))

effects_table
```


```{r}
library(lme4)
library(performance)
library(broom.mixed)
library(purrr)

# Define your dropped blocks from earlier
blocks_dropped <- list(
  semantic = c("cosine_similarity"),
  structure = c("seg1_sentence_length_mean", "seg2_sentence_length_mean"),
  #                     "seg1_n_tokens", "seg2_n_tokens"),
  perplexity = c("seg1_perplexity", "seg2_perplexity", #"seg1_given_seg2_perplexity", 
                 "seg2_given_seg1_perplexity"),
  indirectness = c("seg1_impScore", "seg2_impScore", "prag_distance")
)

# Loop through them and test one-at-a-time addition
block_refit_results_support <- map_df(names(blocks_dropped), function(b) {
  predictors <- paste(blocks_dropped[[b]], collapse = " + ")
  m_block <- update(model_interact_seg1_tree_dep_refined_support, paste(". ~ . +", predictors))
  comp <- anova(model_interact_seg1_tree_dep_refined_support, m_block, test = "Chisq")
  tibble(
    block = b,
    delta_AIC = AIC(model_interact_seg1_tree_dep_refined_support) - AIC(m_block),
    LRT_p = comp$`Pr(>Chisq)`[2],
    R2_marg_gain = performance::r2_nakagawa(m_block)$R2_marginal -
                   performance::r2_nakagawa(model_interact_seg1_tree_dep_refined_support)$R2_marginal
  )
})

block_refit_results_support
```

```{r}
#summary(update(model_interact_three_way_refined, . ~ . + seg1_avg_tree_depth + seg2_avg_tree_depth))
car::Anova(update(model_interact_seg1_tree_dep_refined_support, . ~ . + seg1_perplexity + seg2_perplexity
                  + seg2_given_seg1_perplexity), type="III")
```

```{r}
car::Anova(update(model_interact_seg1_tree_dep_refined_support, . ~ . + seg1_sentence_length_mean
                                                          + seg2_sentence_length_mean), type="III")
```

```{r}
car::Anova(update(model_interact_seg1_tree_dep_refined_support, . ~ . + cosine_similarity), type="III")
```

```{r}
car::Anova(update(model_interact_seg1_tree_dep_refined_support, . ~ . + seg1_impScore
                                                     + seg2_impScore
                                                     + prag_distance), type="III")
```

```{r}
model_final_expanded_support <- glmer(is_support ~ (1|CorpusID) 
                        #+ seg2_n_tokens
                        + seg2_has_deduction
                        + seg1_has_refutation
                        + seg2_has_refutation
                        + seg1_has_condition
                        #+ seg2_has_condition
                        + seg2_has_adverb_of_emphasis
                        + seg1_has_justification
                        + seg2_has_justification
                        #+ seg2_has_modal_verbs
                        + seg2_senti
                        #+ seg1_n_tokens
                        #+ seg1_dependency_distance_mean
                        #+ seg2_dependency_distance_mean
                        + seg2_avg_subclauses *	seg2_dependency_distance_mean
                        + seg1_prop_adjacent_dependency_relation_mean
                        + seg2_prop_adjacent_dependency_relation_mean
                        + seg2_avg_subclauses * seg2_avg_tree_depth
                        + seg1_avg_tree_depth *	seg1_dependency_distance_mean
                        #+ seg1_avg_tree_depth
                        + tfidf_cosine_similarity
                        + seg1_perplexity
                        + seg2_perplexity
                        + seg2_given_seg1_perplexity
                        + cosine_similarity
                        + seg1_impScore
                        + seg2_impScore
                        + prag_distance
                        , data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))

```

```{r}
library(car)
anova(model_interact_seg1_tree_dep_refined_support, model_final_expanded_support, text="Chisq")
print("R2 formerly best")
performance::r2_nakagawa(model_interact_seg1_tree_dep_refined_support)
print("R2 expanded model")
performance::r2_nakagawa(model_final_expanded_support)
car::Anova(model_final_expanded_support, type="III")
performance::check_collinearity(model_final_expanded_support)
```

```{r}
model_final_expanded_refined_support <- glmer(is_support ~ (1|CorpusID) 
                        #+ seg2_n_tokens
                        + seg2_has_deduction
                        + seg1_has_refutation
                        + seg2_has_refutation
                        + seg1_has_condition
                        #+ seg2_has_condition
                        + seg2_has_adverb_of_emphasis
                        + seg1_has_justification
                        + seg2_has_justification
                        #+ seg2_has_modal_verbs
                        + seg2_senti
                        #+ seg1_n_tokens
                        #+ seg1_dependency_distance_mean
                        #+ seg2_dependency_distance_mean
                        + seg2_avg_subclauses *	seg2_dependency_distance_mean
                        #+ seg1_prop_adjacent_dependency_relation_mean
                        + seg2_prop_adjacent_dependency_relation_mean
                        + seg2_avg_subclauses * seg2_avg_tree_depth
                        + seg1_avg_tree_depth *	seg1_dependency_distance_mean
                        #+ seg1_avg_tree_depth
                        + tfidf_cosine_similarity
                        + seg1_perplexity * seg2_given_seg1_perplexity
                        + seg2_perplexity * seg2_given_seg1_perplexity
                        + cosine_similarity
                        + seg1_impScore
                        + seg2_impScore
                        + prag_distance
                        , data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))
```

```{r}
library(car)
anova(model_final_expanded_support, model_final_expanded_refined_support, text="Chisq")

print("R2 expanded model")
performance::r2_nakagawa(model_final_expanded_support)

print("R2 expanded model refined")
performance::r2_nakagawa(model_final_expanded_refined_support)

car::Anova(model_final_expanded_refined_support, type="III")
performance::check_collinearity(model_final_expanded_refined_support)
```

```{r}
library(tidyr)
library(dplyr)
library(tibble)

# 1. Get the fixed-effect formula from your final model
fe_formula <- formula(model_final_expanded_refined_support, fixed.only = TRUE)

# 2. Get all unique, raw variable names from the formula
#    This intelligently extracts "seg1_senti", "cosine_similarity", etc.,
#    and ignores the DV and interaction symbols.
raw_predictors <- all.vars(fe_formula)[-1] # [-1] removes the DV "is_entailment"

# 3. Create a numeric-only dataframe for correlation
model_data_numeric <- hard_threshold_complete %>%
  
  # Select only the predictors that are actually in the final model
  select(all_of(raw_predictors)) %>%
  
  # Convert any factor columns (like "seg1_senti") to numeric 0s and 1s
  # We use as.character(.) first to avoid converting to internal factor levels
  mutate_if(is.factor, as.numeric)

 
cor(model_data_numeric, use = "pairwise.complete.obs")

strong_correlations <- as.data.frame(predictor_cor_matrix) %>%
  rownames_to_column("Var1") %>%
  pivot_longer(
    cols = -Var1,
    names_to = "Var2",
    values_to = "Correlation"
  ) %>%
  filter(abs(Correlation) > 0.4 & abs(Correlation) < 1.0 & Var1 < Var2) %>%
  arrange(desc(abs(Correlation)))

# 4. Print the final, clean table
print(strong_correlations)
```

```{r}
model_final_expanded_remove_per_support <- glmer(is_support ~ (1|CorpusID) 
                        #+ seg2_n_tokens
                        + seg2_has_deduction
                        + seg1_has_refutation
                        + seg2_has_refutation
                        + seg1_has_condition
                        #+ seg2_has_condition
                        + seg2_has_adverb_of_emphasis
                        + seg1_has_justification
                        + seg2_has_justification
                        #+ seg2_has_modal_verbs
                        + seg2_senti
                        #+ seg1_n_tokens
                        #+ seg1_dependency_distance_mean
                        #+ seg2_dependency_distance_mean
                        + seg2_avg_subclauses *	seg2_dependency_distance_mean
                        #+ seg1_prop_adjacent_dependency_relation_mean
                        + seg2_prop_adjacent_dependency_relation_mean
                        + seg2_avg_subclauses * seg2_avg_tree_depth
                        + seg1_avg_tree_depth *	seg1_dependency_distance_mean
                        #+ seg1_avg_tree_depth
                        + tfidf_cosine_similarity
                        + seg1_perplexity 
                        + seg2_given_seg1_perplexity
                        #+ seg2_perplexity * seg2_given_seg1_perplexity
                        + cosine_similarity
                        + seg1_impScore
                        + seg2_impScore
                        + prag_distance
                        , data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))
```

```{r}
library(car)
anova(model_final_expanded_refined_support, model_final_expanded_remove_per_support, text="Chisq")

print("R2 expanded model")
performance::r2_nakagawa(model_final_expanded_refined_support)

print("R2 expanded model remove perplexity")
performance::r2_nakagawa(model_final_expanded_remove_per_support)

car::Anova(model_final_expanded_remove_per_support, type="III")
performance::check_collinearity(model_final_expanded_remove_per_support)
```

```{r}
model_final_expanded_sem_lex_support <- glmer(is_support ~ (1|CorpusID) 
                        #+ seg2_n_tokens
                        + seg2_has_deduction
                        + seg1_has_refutation
                        + seg2_has_refutation
                        + seg1_has_condition
                        #+ seg2_has_condition
                        + seg2_has_adverb_of_emphasis
                        + seg1_has_justification
                        + seg2_has_justification
                        #+ seg2_has_modal_verbs
                        + seg2_senti
                        #+ seg1_n_tokens
                        #+ seg1_dependency_distance_mean
                        #+ seg2_dependency_distance_mean
                        + seg2_avg_subclauses *	seg2_dependency_distance_mean
                        #+ seg1_prop_adjacent_dependency_relation_mean
                        + seg2_prop_adjacent_dependency_relation_mean
                        + seg2_avg_subclauses * seg2_avg_tree_depth
                        + seg1_avg_tree_depth *	seg1_dependency_distance_mean
                        #+ seg1_avg_tree_depth
                        #+ tfidf_cosine_similarity
                        + seg1_perplexity 
                        + seg2_given_seg1_perplexity
                        #+ seg2_perplexity * seg2_given_seg1_perplexity
                        + cosine_similarity * tfidf_cosine_similarity
                        + seg1_impScore
                        + seg2_impScore
                        + prag_distance
                        , data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))
```

```{r}
library(car)
anova(model_final_expanded_refined_support, model_final_expanded_sem_lex_support, text="Chisq")

print("R2 expanded model remove perplexity")
performance::r2_nakagawa(model_final_expanded_refined_support)

print("R2 expanded cos x tfidf")
performance::r2_nakagawa(model_final_expanded_sem_lex_support)

car::Anova(model_final_expanded_sem_lex_support, type="III")
performance::check_collinearity(model_final_expanded_sem_lex_support)
```

```{r}
model_final_expanded_prag_sem_support <- glmer(is_support ~ (1|CorpusID) 
                        #+ seg2_n_tokens
                        + seg2_has_deduction
                        + seg1_has_refutation
                        + seg2_has_refutation
                        + seg1_has_condition
                        #+ seg2_has_condition
                        + seg2_has_adverb_of_emphasis
                        + seg1_has_justification
                        + seg2_has_justification
                        #+ seg2_has_modal_verbs
                        + seg2_senti
                        #+ seg1_n_tokens
                        #+ seg1_dependency_distance_mean
                        #+ seg2_dependency_distance_mean
                        + seg2_avg_subclauses *	seg2_dependency_distance_mean
                        #+ seg1_prop_adjacent_dependency_relation_mean
                        + seg2_prop_adjacent_dependency_relation_mean
                        + seg2_avg_subclauses * seg2_avg_tree_depth
                        + seg1_avg_tree_depth *	seg1_dependency_distance_mean
                        #+ seg1_avg_tree_depth
                        #+ tfidf_cosine_similarity
                        + seg1_perplexity 
                        + seg2_given_seg1_perplexity
                        #+ seg2_perplexity * seg2_given_seg1_perplexity
                        #+ cosine_similarity 
                        + tfidf_cosine_similarity
                        + seg1_impScore
                        + seg2_impScore
                        + prag_distance * cosine_similarity
                        , data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))
```

```{r}
library(car)
anova(model_final_expanded_refined_support, model_final_expanded_prag_sem_support, text="Chisq")

print("R2 expanded model remove perplexity")
performance::r2_nakagawa(model_final_expanded_refined_support)

print("R2 expanded cos x prag_distance")
performance::r2_nakagawa(model_final_expanded_prag_sem_support)

car::Anova(model_final_expanded_prag_sem_support, type="III")
performance::check_collinearity(model_final_expanded_prag_sem_support)
```

```{r}
model_final_expanded_prag_in_support <- glmer(is_support ~ (1|CorpusID) 
                        #+ seg2_n_tokens
                        + seg2_has_deduction
                        + seg1_has_refutation
                        + seg2_has_refutation
                        + seg1_has_condition
                        #+ seg2_has_condition
                        + seg2_has_adverb_of_emphasis
                        + seg1_has_justification
                        + seg2_has_justification
                        #+ seg2_has_modal_verbs
                        + seg2_senti
                        #+ seg1_n_tokens
                        #+ seg1_dependency_distance_mean
                        #+ seg2_dependency_distance_mean
                        + seg2_avg_subclauses *	seg2_dependency_distance_mean
                        #+ seg1_prop_adjacent_dependency_relation_mean
                        + seg2_prop_adjacent_dependency_relation_mean
                        + seg2_avg_subclauses * seg2_avg_tree_depth
                        + seg1_avg_tree_depth *	seg1_dependency_distance_mean
                        #+ seg1_avg_tree_depth
                        #+ tfidf_cosine_similarity
                        + seg1_perplexity 
                        + seg2_given_seg1_perplexity
                        #+ seg2_perplexity * seg2_given_seg1_perplexity
                        #+ cosine_similarity 
                        + tfidf_cosine_similarity
                        + prag_distance * seg1_impScore
                        #+ prag_distance * seg2_impScore
                        + prag_distance * cosine_similarity
                        , data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))
```

```{r}
library(car)
anova(model_final_expanded_prag_sem_support, model_final_expanded_prag_in_support, text="Chisq")

print("R2 expanded cos x prag_distance")
performance::r2_nakagawa(model_final_expanded_prag_sem_support)

print("R2 expanded model prag x impScore")
performance::r2_nakagawa(model_final_expanded_prag_in_support)

car::Anova(model_final_expanded_prag_in_support, type="III")
performance::check_collinearity(model_final_expanded_prag_in_support)
```


```{r}
model_final_support <- model_final_expanded_prag_sem_support

```

```{r}
car::Anova(model_final_support)
```


```{r}
# --- This is your MAIN results table ---
# It's simple, clean, and you already have it.
library(car)
model_significance <- car::Anova(model_final_support, type = "III")
print(model_significance)

# --- This table goes in your APPENDIX ---
library(broom.mixed)
appendix_table <- tidy(model_final_support,
                       effects = "fixed",
                       exponentiate = TRUE,
                       conf.int = TRUE)

print(appendix_table)

# --- This is the ONE-LINE solution ---
library(sjPlot)

# This command creates a beautiful, publication-ready
# HTML table right in your RStudio viewer.
tab_model(
  model_final_support, 
  dv.labels = "Predicting Entailment (is_support = 1)",
  p.style = "scientific"
)
```

```{r}
model_support_simplified <- glmer(is_support ~ (1|CorpusID) 
                        #+ seg2_n_tokens
                        #+ seg2_has_deduction
                        #+ seg1_has_refutation
                        + seg2_has_refutation
                        #+ seg1_has_condition
                        #+ seg2_has_condition
                        #+ seg2_has_adverb_of_emphasis
                        #+ seg1_has_justification
                        #+ seg2_has_justification
                        #+ seg2_has_modal_verbs
                        + seg2_senti
                        #+ seg1_n_tokens
                        #+ seg1_dependency_distance_mean
                        #+ seg2_dependency_distance_mean
                        #+ seg2_avg_subclauses *	seg2_dependency_distance_mean
                        #+ seg1_prop_adjacent_dependency_relation_mean
                        #+ seg2_prop_adjacent_dependency_relation_mean
                        #+ seg2_avg_subclauses * seg2_avg_tree_depth
                        #+ seg1_avg_tree_depth *	seg1_dependency_distance_mean
                        #+ seg1_avg_tree_depth
                        #+ tfidf_cosine_similarity
                        #+ seg1_perplexity 
                        #+ seg2_given_seg1_perplexity
                        #+ seg2_perplexity * seg2_given_seg1_perplexity
                        #+ cosine_similarity 
                        + tfidf_cosine_similarity
                        + seg1_impScore
                        #+ seg2_impScore
                        #+ prag_distance * cosine_similarity
                        , data = hard_threshold_complete, family = binomial, control = glmerControl(optimizer = "bobyqa"))

```

```{r}
library(car)
anova(model_final_support, model_support_simplified, text="Chisq")
performance::r2_nakagawa(model_final_support)
performance::r2_nakagawa(model_support_simplified)
car::Anova(model_support_simplified, type="III")
performance::check_collinearity(model_support_simplified)
```